---
title: "use_cases"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{use_cases}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(iedbr)
```

```{r good_to_go}
good_to_go <- beastier::is_on_ci()
```



In healthy humans, 

 * Get all epitopes from all B cells

```{r}
if (good_to_go) {
  epitopes <- get_all_b_cell_epitopes() 
  head(epitopes)
}
```

 * Get all epitopes from B cells of a certain MHC allele

```
if (good_to_go) {
  epitopes <- get_all_b_cell_epitopes(haplotype = ) 
  head(epitopes)
}
```

 * Get all epitopes from all T cells
 
```{r}
if (good_to_go) {
  epitopes <- get_all_t_cell_epitopes() 
  head(epitopes)
}
```

 * Get all epitopes from T cells of a certain MHC allele
 * Get all epitopes from all MHC ligand assays

```{r}
if (good_to_go) {
  epitopes <- get_all_mhc_ligand_epitopes() 
  head(epitopes)
}
```

 * Get all epitopes from MHC ligand assays of a certain MHC allele

```
require(httr)

args <- commandArgs(trailingOnly = TRUE)

if (1 == 2) {
  setwd("~/GitHubs/bbbq_article_issue_157")
  list.files()
  args <- c("iedb_b_cell", "per_allele", 2)
  args <- c("iedb_t_cell", "per_allele", 1)
  args <- c("iedb_mhc_ligand", "all_alleles", 1)
  args <- c("iedb_mhc_ligand", "all_alleles", 2)
  args <- c("iedb_b_cell", "all_alleles", 1)
  args <- c("iedb_mhc_ligand", "per_allele", 2)
}
message("args: {", paste0(args, collapse = ", "), "}")
testthat::expect_equal(length(args), 3)
dataset <- as.character(args[1])
message("dataset: ", dataset)
testthat::expect_true(
  dataset %in% c(
    "schellens", "bergseng", "iedb_b_cell", "iedb_t_cell","iedb_mhc_ligand"
  )
)
allele_set <- as.character(args[2])
testthat::expect_true(allele_set %in% c("per_allele", "all_alleles"))
message("allele_set: ", allele_set)

mhc_class <- as.numeric(args[3])
message("mhc_class: ", mhc_class)
testthat::expect_true(mhc_class %in% c(1, 2))

output_filename <- paste0(dataset, "_", allele_set, "_", mhc_class, ".csv")
message("output_filename: ", output_filename)

which_cells <- NA
if (dataset == "iedb_t_cell") which_cells <- "t_cells"
if (dataset == "iedb_b_cell") which_cells <- "b_cells"
if (dataset == "iedb_mhc_ligand") which_cells <- "mhc_ligands"
testthat::expect_true(!is.na(which_cells))
message("which_cells: ", which_cells)


  
tibbles <- list()
i <- 1
haplotypes <- NA
if (mhc_class == 1) haplotypes <- bbbq::get_mhc1_haplotypes()
if (mhc_class == 2) haplotypes <- bbbq::get_mhc2_haplotypes()
if (allele_set == "all_alleles") haplotypes <- "all"
n_haplotypes <- length(haplotypes)

for (haplotype in haplotypes) {
  # Use the IEDB names
  # Don't need in new BBBQ version
  haplotype <- stringr::str_replace_all(
    haplotype, "\\*([[:digit:]]{2})([[:digit:]]{2})", 
    "*\\1:\\2"
  )
  message(
    i, "/", n_haplotypes, ": ",
    "haplotype: ", haplotype, ", ",
    "which_cells: ", which_cells, ", ",
    "mhc_class: ", mhc_class
  )
  if (which_cells == "b_cells") {
    params <- list(
      `structure_type` = 'eq.Linear peptide',
      `host_organism_iris` = 'cs.{NCBITaxon:9606}',
      `source_organism_iris` = 'cs.{NCBITaxon:9606}',
      `disease_names` = 'cs.{healthy}',
      `order` = 'structure_iri'
    )
    if (allele_set == "per_allele") {
      params$mhc_allele_names <- paste0("cs.{", haplotype, "}")
    } else {
      # Will filter out the correct MHC class later
    }
    params$bcell_ids <- 'not.is.null'
    res <- httr::GET(url = 'https://query-api.iedb.org/epitope_search', query = params)
  } else if (which_cells == "t_cells") {
    params <- list(
      `structure_type` = 'eq.Linear peptide',
      `host_organism_iris` = 'cs.{NCBITaxon:9606}',
      `source_organism_iris` = 'cs.{NCBITaxon:9606}',
      `disease_names` = 'cs.{healthy}',
      `order` = 'structure_iri'
    )
    if (allele_set == "per_allele") {
      params$mhc_allele_names <- paste0("cs.{", haplotype, "}")
    } else {
      # Will filter out the correct MHC class later
    }
    params$tcell_ids <- 'not.is.null'
    res <- httr::GET(url = 'https://query-api.iedb.org/epitope_search', query = params)
  } else {
    testthat::expect_equal(which_cells, "mhc_ligands")
    params <- list(
      `structure_type` = 'eq.Linear peptide',
      `disease_names` = 'cs.{healthy}',
      `order` = 'structure_iri'
    )
    res <- httr::GET(url = 'https://query-api.iedb.org/mhc_search', query = params)
  }
  content <- httr::content(res)
  message("Found ", length(content), " candidates")
  if ("message" %in% names(content)) {
    if (
        stringr::str_detect(
        string = content$message,
        pattern = "column .* does not exist"
      )
    ) {
      stop("Does not exist")
    }
  }
  if (!is.list(content)) stop("'content' must be a list")
  if (length(content) == 0) {
    message("No results for haplotype ", haplotype)
    next
  }
  testthat::expect_true("linear_sequence" %in% names(content[[1]]))
  linear_sequences <- purrr::map_chr(content, function(x) { x$linear_sequence } )
  t <- tibble::tibble(linear_sequence = linear_sequences)
  if (which_cells == "b_cells" || which_cells == "t_cells")  {
    # Do not check for MHC binding assays
    # content[[1]]
    # are_mhc_binding_essays <- purrr::map_lgl(content, function(x) { "MHC binding assay" %in% x$mhc_allele_evidences } ) 
    # head(are_mhc_binding_essays)
    # testthat::expect_equal(length(linear_sequences), length(are_mhc_binding_essays))
    # t <- t[are_mhc_binding_essays, ]
    # testthat::expect_equal(nrow(t), sum(are_mhc_binding_essays))
  } else {
    testthat::expect_equal(which_cells, "mhc_ligands")
    if (allele_set == "per_allele") {
      is_correct_haplotype <- purrr::map_lgl(content, function(x) {  haplotype == x$mhc_allele_name } ) 
      message("Candidates with correct haplotype: ", sum(is_correct_haplotype))
      t <- t[is_correct_haplotype, ]
      testthat::expect_equal(nrow(t), sum(is_correct_haplotype))
    } else {
      testthat::expect_equal(allele_set, "all_alleles")
      # content[[1]]$mhc_class
      # as.character(as.roman(mhc_class)) == content[[1]]$mhc_class
      is_correct_mhc_class <- purrr::map_lgl(content, function(x) {  as.character(as.roman(mhc_class)) == x$mhc_class } ) 
      t <- t[is_correct_mhc_class, ]
      testthat::expect_equal(nrow(t), sum(is_correct_mhc_class))
    }
  }
  t <- dplyr::distinct(t)
  t$haplotype <- haplotype 
  t$cell_type <- which_cells 
  message("Got ", nrow(t), " new epitopes")
  tibbles[[i]] <- t
  i <- i + 1
}
t <- dplyr::bind_rows(tibbles)
t
readr::write_csv(t, output_filename)
testthat::expect_true(file.exists(output_filename))
```



# Multiple quesies

From https://github.com/IEDB/IQ-API-use-cases/blob/master/R/uniprot_reference_proteins.Rmd:

Let's set the query string parameters.

```
params <- list()
params[['parent_source_antigen_iri']] = 'like.UNIPROT*'
params[['select']] = 'parent_source_antigen_id,parent_source_antigen_iri,parent_source_antigen_source_org_iri,parent_source_antigen_name'
params[['offset']] = 0
params[['order']] = 'parent_source_antigen_id'
```

Since I know there will be more than 10K records, we will need to fetch 1 page at a time
and continue to append them onto our tibble.  **NOTE:** When paging through results by using an 'offset', it is critical to add an 'order' parameter to ensure that the pages are consistent between queries.

To be good citizens, we also add a short 'sleep' in between calls to the API.
```
# initialize 'get_text' so we can page through the results
get_text <- 'NA'
full_tbl <- tibble()
while(get_text != '[]') {
  url <- paste(base_uri, 'antigen_search',sep='')
  print(paste0("fetching offset: ",params[['offset']]))
  get_1 = GET(url, query=params)
  get_text = content(get_1,'text')
  resp_tbl <- tibble(fromJSON(get_text))
  full_tbl <- rbind(full_tbl, resp_tbl)
  params[['offset']] = params[['offset']] + 10000
  # sleep for 1 second between calls so as not to overload the server
  Sys.sleep(1)
}
```
